\documentclass[10pt, letterpaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[letterpaper, top=0.6in, bottom=0.6in, left=0.6in, right=0.6in]{geometry}
\usepackage{enumitem}
\usepackage{titlesec}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{charter}
\usepackage{microtype}

% Color Definitions - Sleek Dark Navy & Slate
\definecolor{primary}{RGB}{20, 50, 110} 
\definecolor{text}{RGB}{30, 30, 30}
\definecolor{linkcolor}{RGB}{20, 50, 110}

\hypersetup{
    colorlinks=true,
    linkcolor=linkcolor,
    filecolor=linkcolor,
    urlcolor=linkcolor,
    pdfauthor={Zerihun Ayele Chala},
    pdftitle={Zerihun Ayele Chala - Resume}
}

% Clean Academic Formatting
\titleformat{\section}{\large\bfseries\color{primary}\uppercase}{}{0pt}{}[\titlerule]
\titlespacing{\section}{0pt}{12pt}{6pt}

% Custom list spacing for tight, professional layout
\setlist[itemize]{leftmargin=0.15in, itemsep=2.5pt, parsep=1pt, topsep=2pt, partopsep=0pt}

\begin{document}

% --- HEADER ---
\begin{center}
    {\Huge \textbf{\color{primary} Zerihun Ayele Chala}} \\
    \vspace{6pt}
    \small
    Addis Ababa, Ethiopia \ | \ 
    \href{mailto:zachrist80@gmail.com}{zachrist80@gmail.com} \ | \ 
    \href{https://www.linkedin.com/in/zerihun-ayele-chala}{linkedin.com/in/zerihun-ayele-chala}
\end{center}

\vspace{-4pt}

% --- PROFESSIONAL SUMMARY ---
\section{Professional Summary}
Computational researcher and deep learning engineer with a rigorous mathematical foundation in dynamical systems, statistical mechanics, and high-dimensional optimization. Specializes in framing deep architectures through a statistical lens—interpreting neural networks as continuous-time dynamical systems, analyzing gradient flow via spectral properties of Jacobians/Hessians, and modeling learning dynamics under probabilistic constraints. Expert in Scientific Machine Learning (SciML) and high-performance computation (C++, JAX, CUDA, PyTorch), with a proven track record of bridging continuum mechanics, control theory, and biophysical neuroscience with modern generative and sequence modeling paradigms.

% --- EDUCATION ---
\section{Education}
\textbf{Addis Ababa University} \hfill \textbf{Oct 2025 -- Present} \\
\textit{MSc in Artificial Intelligence} \hfill \textbf{Cumulative Status: Year 1 Completed (Jun 2026)} \\
Research focus: Mathematical modeling of neural dynamics, Scientific Machine Learning (SciML), and Neural ODEs.

\vspace{6pt}
\textbf{Addis Ababa University} \hfill \textbf{Sep 2015 -- Dec 2020} \\
\textit{BSc in Mechanical Engineering} \hfill \textbf{University Gold Medalist (Ranked 1st/Cohort)} \\
\textbf{GPA: 3.83/4.00 (Very Great Distinction)}. A 5-year rigorous curriculum establishing a deep foundation in Multivariable Calculus, Linear Algebra, Complex Analysis, and Non-linear Differential Equations. Advanced focus on Continuum Mechanics, Fluid Dynamics (Navier-Stokes), and Control Theory.

% --- TECHNICAL & SCIENTIFIC EXPERTISE ---
\section{Computational \& Mathematical Mastery}
\begin{itemize}
    \item \textbf{Mathematical Foundations:} Dynamical Systems, Stochastic Calculus, PDE/ODE Solvers, Jacobian/Hessian Spectral Analysis, Convex \& Non-Convex Optimization, Statistical Mechanics, and Information Theory.
    \item \textbf{Deep Learning \& Statistical Inference:} Generative Modeling (Diffusion, VAEs, GANs), Sequence Models (State-Space Models, Transformers), Physics-Informed Neural Networks (PINNs), Variational Inference, and Bayesian Neural Networks.
    \item \textbf{Programming \& Systems:} Advanced C++ (C++17/20, OOP, memory management, multithreading), Python (NumPy, SciPy, PyTorch, JAX, Equinox), CUDA, MATLAB/Simulink, Git, and Linux environments.
    \item \textbf{Biophysical Neuroscience:} Hodgkin-Huxley model, Synaptic Gating Kinetics, Cable Theory, Synaptic Plasticity (STDP, LTP/LTD), and Computational Modeling of Cortical Circuits.
\end{itemize}

% --- ACADEMIC EXPERIENCE ---
\section{Academic Leadership}
\textbf{Mizan-Tepi University} \hfill \textbf{Mar 2021 -- Feb 2025} \\
\textit{Lecturer in Computational Engineering \& Numerical Methods}
\begin{itemize}
    \item Instructed 100+ undergraduate students on numerical methods, focusing on the translation of high-dimensional calculus, numerical linear algebra, and differential equations into optimized, executable code in C++ and Python.
    \item Taught core courses in Algorithms \& Data Structures, emphasizing memory management, computational complexity (Big-O analysis), and cache-efficient algorithm design for scientific computing.
    \item Developed rigorous curriculum and hands-on laboratory exercises for numerical integration (Runge-Kutta, Gaussian quadrature) and matrix decomposition algorithms (LU, QR, Cholesky, and SVD).
\end{itemize}

\newpage

% --- RESEARCH & ALGORITHMIC IMPLEMENTATIONS ---
\section{Selected Implementations \& Research Projects}

\textbf{High-Performance C++/CUDA Tensor \& Automatic Differentiation Engine} 
\begin{itemize}
    \item Engineered a modular, high-performance C++/CUDA tensor library and reverse-mode automatic differentiation engine from scratch without relying on high-level deep learning frameworks.
    \item Designed an execution graph structure (DAG) with tape-based gradient tracking and implemented custom memory pools to optimize tensor reuse, reducing GPU memory fragmentation.
    \item Wrote and optimized custom CUDA kernels for element-wise operations, activations, and tiled General Matrix Multiplication (GEMM) using shared memory, achieving execution speeds matching standard libraries for small-to-medium batch sizes.
    \item Exposed C++ internals to Python using PyBind11, enabling seamless prototyping and validation against PyTorch.
\end{itemize}

\textbf{Scientific Machine Learning: Physics-Informed Fourier Neural Operators (PINO) for Fluid Dynamics}
\begin{itemize}
    \item Engineered a Physics-Informed Fourier Neural Operator (PINO) in JAX to solve the 2D/3D Navier-Stokes equations for incompressible fluid flows under complex boundary conditions.
    \item Vectorized continuous PDE residuals using automatic differentiation, incorporating them directly into the loss function to act as a soft physical constraint.
    \item Achieved up to 100x acceleration in simulation speed compared to classical Finite Element Method (FEM) solvers, while maintaining physical consistency (mass and momentum conservation) within a 0.5\% error margin.
    \item Leveraged parallel JAX scans and multi-GPU compilation (\texttt{jit}/\texttt{pmap}) to handle high-dimensional spatial grids.
\end{itemize}

\textbf{Stochastic Generative Modeling: Score-Based Diffusion Models via Continuous-Time SDEs}
\begin{itemize}
    \item Developed a Denoising Diffusion Probabilistic Model (DDPM) framed through continuous-time Stochastic Differential Equations (SDEs), utilizing the Variance Preserving (VP) and Variance Exploding (VE) formulations.
    \item Trained score networks in PyTorch to estimate the score function of high-dimensional data distributions, utilizing sliced score matching and denoising score matching.
    \item Implemented numerical solvers for the reverse-time Ito SDE (Euler-Maruyama) and the deterministic Probability Flow ODE (Runge-Kutta 4th Order), enabling exact likelihood computation and high-fidelity sampling.
    \item Investigated the connection between stochastic dynamics, entropy production, and sampling efficiency across different temperature schedules.
\end{itemize}

\textbf{Sequence Modeling \& Control: Hardware-Aware Structured State-Space Models (S4/Mamba) from Scratch}
\begin{itemize}
    \item Implemented a Structured State-Space (S4) sequence model in JAX, establishing an explicit mathematical mapping between classical control theory state-space representations and modern sequence learning.
    \item Designed and optimized a parallel associative scan in JAX to bypass sequential recurrence bottlenecks, achieving $O(\log N)$ parallel computation over sequences of length $N$.
    \item Formulated discrete-time state transitions from continuous-time matrices via bilinear (Tustin) and Zero-Order Hold (ZOH) transformations, analyzing their stability and spectral properties.
    \item Validated the model on the Long Range Arena (LRA) benchmark, demonstrating superior memory capacity and linear scaling relative to standard self-attention mechanisms.
\end{itemize}

% --- SELECTED HONORS & CERTIFICATIONS ---
\section{Selected Honors \& Certifications}
\begin{itemize}
    \item \textbf{Synapses, Neurons and Brains (Hebrew University of Jerusalem):} Intensive specialization in the biophysics of neurons, synaptic transmission, and the computational logic of neural circuits.
    \item \textbf{University Gold Medal (2020):} Highest academic honor, awarded to the top-ranking student in the Mechanical Engineering Department (Ranked 1st out of cohort).
    \item \textbf{Deep Learning Specialization (DeepLearning.AI):} Comprehensive mastery of Convolutional Neural Networks (CNNs), Recurrent Neural Networks (RNNs), and Sequence Models.
    \item \textbf{Distinguished Lecturer Award (2023):} Recognized for excellence, clarity, and innovation in teaching technical computational engineering subjects.
\end{itemize}

\end{document}
